---
title: "Covid-19 in Madagascar"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    social: [ "twitter", "facebook", "menu" ]
    vertical_layout: scroll
    theme: bootstrap
    navbar:
      - { title: "<img style=\"width: 16px;\" src=\"www/us16.png\" />   English", href: "index.html"}
     
params:
  lang: "EN"
---

<!-- Google Tag Manager -->
<script>(function(w,d,s,l,i){w[l]=w[l]||[];w[l].push({'gtm.start':
new Date().getTime(),event:'gtm.js'});var f=d.getElementsByTagName(s)[0],
j=d.createElement(s),dl=l!='dataLayer'?'&l='+l:'';j.async=true;j.src=
'https://www.googletagmanager.com/gtm.js?id='+i+dl;f.parentNode.insertBefore(j,f);
})(window,document,'script','dataLayer','GTM-N9D53DP');</script>
<!-- End Google Tag Manager -->

```{r setup, include=FALSE}
library(flexdashboard)
library(googlesheets4)
library(shiny)
library(sf)
library(leaflet)
library(tidyverse)
library(plotly)
library(lubridate)
library(dygraphs)
library(zoo)
library(forecast)
library(stringdist)
library(glue)
```


```{r}

# pull data from google sheets & clean
gs4_deauth()
data <- read_sheet("https://docs.google.com/spreadsheets/d/1oQJl4HiTviKAAhCjMmg0ipGcO79cZg6gSHrdTuQID_w/edit#gid=0", sheet = 2)
test <- read_sheet("https://docs.google.com/spreadsheets/d/1oQJl4HiTviKAAhCjMmg0ipGcO79cZg6gSHrdTuQID_w/edit#gid=0", sheet = 3)
events <- read_sheet("https://docs.google.com/spreadsheets/d/1oQJl4HiTviKAAhCjMmg0ipGcO79cZg6gSHrdTuQID_w/edit#gid=0", sheet = 5)
text <- read_sheet("https://docs.google.com/spreadsheets/d/14N9y4A9MLeYeNyAIdWRwluwhSxQdBi7DT_2CDuf_LUg/edit#gid=0")

# shapefiles
mdg2 <- read_sf("data/regions/MDG_ADM1.shp")

# Cases by region
data$region<-mdg2$NAME[amatch(data$Location4,mdg2$NAME,maxDist=6)]
data$Date<-as.Date(data$Date)
data$region[is.na(data$region)]="unspecified"
data %>%
  group_by(region) %>%
  summarize(cases = sum(Type=="N"), deaths = sum(Type == "D")) -> cases_by_region


deaths<-filter(data,Type=="D")
cases<-filter(data,Type=="N")

cases_by_date <- as.data.frame(xtabs( ~ Date+region , cases))
cases_by_date$Date<-as.Date(cases_by_date$Date)
cases_by_date%>%rename(cases=Freq)%>%
  group_by(region)%>%
  mutate(cum.sum=cumsum(cases))%>%#compute cumulative number of cases per region
  mutate(mov.avg=rollmeanr(cases,7,fill="extend"))->cases_by_date#compute 7 days moving average per region

deaths$Age<-as.numeric(deaths$Age)
deaths_by_date<-as.data.frame(xtabs( ~ Date+region , deaths))
deaths_by_date$Date<-as.Date(deaths_by_date$Date)
deaths_by_date%>%
  rename(deaths=Freq)%>%
  group_by(region)%>%
  mutate(deaths.cum.sum=cumsum(deaths))->deaths_by_date

#IRD for infected (new cases), Recovered (R) and Removed (D)
IRD<-as.data.frame(xtabs( ~ data$Date+data$Type))
IRD%>%rename("Date"=1,"Type"=2,"n"=3)%>%filter(Type!="S")->IRD
IRD$Date<-as.Date(IRD$Date)

IRD%>%
  arrange(Date)%>%
  dplyr::group_by(Type)%>%
  mutate(Type.cum.sum=cumsum(n))->IRD


##testing

test %>%select("Date"=Date,"Cum.cases"=Cum.cases,"New.cases"=new.cases, "Total.tests"=Global.total.test)%>%arrange(Date)->test

Diff.previous<-data.frame("Date"=test$Date[c(2:length(test$Total.tests))],
                          "tests.diff"=diff(test$Total.tests, lag = 1,na.rm=T),                                        "time.diff"=as.integer(diff.Date(test$Date)))%>%
  filter(!is.na(tests.diff)&time.diff==1)#table of daily test (when  there is exactly 1 day between two available "numbers of tests")

data%>%
  group_by(Date)%>%
  summarise(n.cases=sum(Type=="N"),n.deaths=sum(Type=="D"))%>%
  arrange(Date)%>%
  mutate(cum.cases=cumsum(n.cases),mov.avg=rollmeanr(n.cases,7,fill="extend"),
         cum.deaths=cumsum(n.deaths),d.mov.avg=rollmeanr(n.deaths,7,fill="extend"))%>%
  right_join(mutate(test, Date = ymd(Date))) %>%
  left_join(mutate(Diff.previous, Date = ymd(Date))) %>%
  arrange(Date)%>%
  mutate(tests.mov.avg=rollmeanr(tests.diff,5,fill="extend"))%>%
  mutate(positivity=(n.cases/tests.diff)*100)%>%
  mutate(positivity.mov.avg=rollmeanr(positivity,5,fill="extend"))%>%
  select(Date,n.cases,cum.cases,mov.avg,n.deaths,cum.deaths,d.mov.avg,n.tests=tests.diff,cum.tests=Total.tests,tests.mov.avg,positivity,positivity.mov.avg)-> ntl.cases


ntl.cases<-ntl.cases[order(ntl.cases$Date,na.last=FALSE),]
ntl.cases$cum.cases<-na.locf(ntl.cases$cum.cases)
ntl.cases$cum.deaths<-na.locf(ntl.cases$cum.deaths)
ntl.cases$Date<-as.Date(ntl.cases$Date)
```

``` {r translation}
# Set up language for rest of dashboard
text <- mutate(text, lang = get(params$lang))

# helper functions to pull correct language (defaults to English when no available translation)

# larger chunks of text by a reference
translate <- function(ref) {
  txt <- text$lang[text$ref %in% ref]
  if (identical(txt, character(0)) | is.na(txt)) {
    glue(text$EN[text$ref %in% ref])
  } else {
    glue(txt)
  }
}

# single line by the actual text in English
translate.sl <- function(sl) {
  txt <- text$lang[text$EN %in% sl]
  if (identical(txt, character(0)) | is.na(txt)) {
    glue(text$EN[text$EN %in% sl])
  } else {
    glue(txt)
  }
}

# Notes on translation:
# For longer chunks you can use the ref column to match your text
# For single line labels etc, you can just search for the English text (see the value boxes as an example!)
# See the google sheet linked here for translations, write the translations using Rmd syntax!
```

Country profile: Madagascar
========================================================================

Sidebar {.sidebar}
-----------------------------------------------------------------------

```{r sidebar, results='asis', verbatim=TRUE}

date_updated <- format(last(cases$Date),format="%B %d %Y") # any inline stats you need ahead of time!
translate("sidebar")

```


Row 
-----------------------------------------------------------------------

### Total cases

```{r}
valueBox(value=format(sum(cases_by_date$cases),big.mark = ","),
         caption=paste(translate.sl("Number of cases reported in Madagascar since 20-March-2020 until"), 
                            format(last(cases_by_date$Date), format="%B %d %Y")),
         icon="fa-ambulance",color="info")
```

### Daily cases

```{r}
valueBox(value=format(last(ntl.cases$mov.avg),big.mark = ",",digits=3),
         caption=paste(translate.sl("Daily new cases (7 days moving average)"), format(last(cases_by_date$Date), format="%B %d %Y")),
         icon="fa-chart-line", color="danger")
```

### Recovered cases
```{r}
IRD%>%filter(Type=="R")->recov
valueBox(value=format(last(recov$Type.cum.sum),big.mark = ","),
         caption=paste(translate.sl("Number of recoveries reported in Madagascar since 20-March-2020 until"), format(last(cases_by_date$Date), format(last(recov$Date), format="%B %d %Y"))), icon="fa-check-circle",color="success")
```

### Total deaths
```{r}
valueBox(value=sum(cases_by_region$deaths,na.rm=TRUE), 
         caption=translate.sl("Total Number of deaths caused by COVID-19 reported in Madagascar since 20 March 2020"),color="danger",icon="fa-heartbeat",href="### Total deaths")
```


### Total tests performed: How many tests did Madagascar so far?
```{r}
valueBox(value=format(last(test$Total.tests),big.mark = ","),caption=paste(translate.sl("Total Number of tests performed in Madagascar between 20-March-2020 and"), format(last(test$Date),format= "%B %d %Y")),
         icon="fa-flask",color=ifelse(last(test$Total.tests)/27692<5,"warning","info"))
```

Row
------------------------

```{r notes, results='asis', verbatim=TRUE}

translate("notes")

```


Row {data-width=650}
-----------------------------------------------------------------------

### **Map of cases**: How many cases have been detected in each regions of Madagascar. 

The press briefings failed to report the origin of some cases. These omissions will affect the accuracy of the figures below

```{r}
mada_regions <- left_join(mdg2, cases_by_region, by = c("NAME" = "region"))

bins <- pretty(cases_by_region$cases,n=12)
pal <- colorBin("YlOrRd", domain = mada_regions$cases)

labels <- sprintf(
      "<strong>%s</strong><br/> Cases: %i <br/> Deaths: %i",
      mada_regions$NAME, mada_regions$cases, mada_regions$deaths
    ) %>% lapply(htmltools::HTML)
    
map <-leaflet()  %>%
  addTiles()%>%
  addPolygons(data = mada_regions,
              color = "black", weight = 0.001, smoothFactor = 0.1,
              fillColor = ~pal(cases),
              fillOpacity = 0.6,
              dashArray = NULL,
              label = labels,
              highlightOptions = highlightOptions(
                    weight = 3,
                    color = "black",
                    dashArray = NULL,
                    fillOpacity = 0.75,
                    bringToFront = TRUE),
              labelOptions = labelOptions(
                style = list("font-weight" = "normal", padding = "3px 8px"),
                textsize = "15px",
                direction = "auto")) %>%
  addLegend("bottomright", pal = pal, values = bins, title = "Number of reported cases")
map %>%addProviderTiles(providers$Esri.WorldStreetMap)
```

### **COVID 19 in Madagascar?**
COVID-19 is the disease caused by the novel Coronavirus (SARS-CoV2) that emerged and was first detected in Wuhan, China in November/December 2019. 

Through human movement (migration and travel) the virus has spread quickly throughout the World and has infected over 10 million people and killed more than 500 000 of those (July 2, 2020).

In response to the outbreak in China, the Government of Madagascar suspended flights to and from China on February 10th and then all international flights were suspended as of March 20, 2020.

Nevertheless, COVID 19 was first detected in Madagascar on **March 20, 2020** from three individuals with recent travel history from Europe. 

Authorities declared a Public Health Emergency through which, confinement measures were set in order to limit movement within and between regions of the Island. Some of these measures have since been relaxed.

Testing is available in the country mainly for suspected cases. The [Institut Pasteur de Madagascar](www.pasteur.mg) acquired the capabilities to conduct the tests in January 2020 and has conducted more than 14 000 tests so far. Recently, three other laboratories acquired the capabilities to conduct the tests too. Testing laboratories include [Institut pasteur de Madagascar](www.pasteur.mg), [the Centre d'Infectiologie Charles Merieux](https://cicm-madagascar.com), the HJRA hospital and the Laboratoire d'Analyses Medicales de Madagascar.  Collectively, these laboratories conduct **on average `r format(round(last(na.trim(ntl.cases$tests.mov.avg)),2),big.mark = ",")` every day  (5 day moving average)**.

As of **`r format(last(cases_by_date$Date),format="%B %d %Y")`**, Madagascar has confirmed **`r format(sum(cases_by_date$cases),big.mark = ",")`** cases in `r n_distinct(data$Location4,na.rm=TRUE)` regions and `r length(deaths$Id)` deaths.


Row {data-width=650}
-----------------------------------------------------------------------


### **Time series: new cases, recoveries and deaths**

```{r}
ggplot(data=IRD,aes(x=Date,y=Type.cum.sum,fill=Type))+
  geom_area(data=IRD,aes(x=Date,y=Type.cum.sum,fill=Type),alpha=0.7,size=2,position="dodge")+
  scale_fill_discrete(labels=c("New cases","Recovered","Deaths"),breaks=c("N","R","D"))+
  scale_x_date(date_breaks = "2 weeks")+
  theme(panel.background=element_blank(),axis.text.x = element_text(angle = 45, hjust = 1))+
  labs(y="Number of people")->IRD_curve
ggplotly(IRD_curve)

```


### **Cumulative number of cases**: How many cases were detected in each region through time?


```{r}
library(ggthemes)
library(scales)
library(RColorBrewer)

colourCount =length(unique(cases_by_date$region))
mycolors <- colorRampPalette(brewer.pal(8, "Paired"))(colourCount)
ggplot(data=cases_by_date,aes(x=Date,y=cum.sum,fill=region))+
  geom_area(data=cases_by_date,aes(x=Date,y=cum.sum,fill=region),size=0.75)+
  scale_fill_manual(values=mycolors)+scale_x_date(date_breaks = "2 weeks")+theme(panel.background=element_blank(),axis.text.x = element_text(angle = 45, hjust = 1))+labs(y="Number of cases")->cumulative_cases
ggplotly(cumulative_cases)

```


Row {data-width=650}
-------------------------------

### **Daily number of cases and moving average**

```{r, warning=FALSE}
daily.cases<-ggplot(data=ntl.cases,aes(x=Date,y=mov.avg))+
  geom_col(aes(x=Date,y=n.cases,fill="n.cases"),position="dodge",alpha=0.4)+
  scale_x_date(date_breaks = "2 weeks") +
  geom_line(aes(x=Date,y=mov.avg,color="mov.avg"),size=1)+
  labs(y="Daily number of cases")+
  scale_fill_manual(name = "", values = c("n.cases" = "#1B9E77")) +
  scale_color_manual(name= "",values = c("mov.avg" = "#E6AB02")) +
  theme(panel.background=element_blank(),axis.text.x = element_text(angle = 45, hjust = 1))
ggplotly(daily.cases)

```


Row {data-width=650}
----------------------------------

### **Daily Number of cases per region** 
```{r cases_movavg, warning=FALSE}
modavg.curv<-ggplot(data=ungroup(cases_by_date, region) ,aes(x=Date,y=mov.avg))+
  geom_col(aes(x=Date,y=cases,fill=region),position="dodge",alpha=0.6)+
  scale_x_date(date_breaks = "2 weeks") +
  geom_smooth(aes(x=Date,y=mov.avg,group=region,colour=region),size=0.7,method="gam")+
  labs(y="Daily number of cases")+
  theme(panel.background=element_blank(),axis.text.x = element_text(angle = 45, hjust = 1))
ggplotly(modavg.curv)


```


### **Selected Metrics and indicators**

**Test**: The tests used to confirm that a person is infected with the virus is called a RT-PCR (Real time Polymerase Chain Reaction). It allows scientists detect genetic material that belongs to the virus and suggest that the person is **currently** infected (can pass the virus to another person). 


**Case**: An individual who obtained a positive result after a molecular test by RT-PCR from one of the approved laboratories in the country, whether or not the patient presented symptoms to the disease.
The reported case figures on a given date does not necessarily show the number of new cases on that day: this is due to delays in testing and reporting.

**Moving average**: The 5 or 7 day moving average is calculated everyday and consists of the average of the values of an indicator (e.g. cases) during the previous 5 or 7 days respectively. It allows to evaluate the overall trend of this indicator and helps prevent major events from skewing the data.

**Positivity rate**: This corresponds to the proportion of confirmed cases per 100 tests performed. The WHO has suggested a positivity rate of around 3–12% as a general benchmark of adequate testing. [reference](https://www.who.int/docs/default-source/coronaviruse/transcripts/who-audio-emergencies-coronavirus-press-conference-full-30mar2020.pdf?sfvrsn=6b68bc4a_2)

**Case fatality rate**: The proportion of confirmed cases that died of COVID-19 so far. 


Row 
-----------------------------------------------------------------------

### Total tests performed: How many tests did Madagascar perform so far?
```{r}
valueBox(value=format(last(test$Total.tests),big.mark = ","),caption=paste("Total Number of tests performed in Madagascar between 20-March-2020 and",format(last(test$Date),format= "%B %d %Y")),
         icon="fa-flask",color=ifelse(last(test$Total.tests)/27692<5,"warning","info"))
```

### Tests per thousand: How many tests did Madagascar perform so far?
```{r}
valueBox(value=format(round(last(test$Total.tests/27692),2),big.mark = ","),caption="Tests per thousand habitants performed in Madagascar",
         icon="fa-vial",color=ifelse(last(test$Total.tests)/27.692<5,"warning","info"))
```

### Positivity (5-day moving average)

```{r}
positivity_ma<-ntl.cases%>%filter(!is.na(positivity.mov.avg))
gauge(round(last(positivity_ma$positivity.mov.avg),2),min=0,max=100,symbol='%',label="Positivity rate",abbreviateDecimals = 2,
      sectors=gaugeSectors(success=c(0,5),warning = c(5,20),danger=c(20,100)))
```



### Daily tests (moving average)
```{r}

valueBox(value=format(round(last(na.trim(ntl.cases$tests.mov.avg)),2),big.mark = ","), caption="Daily number of tests (5 days moving average)",
         icon="fa-vial",color=ifelse(ntl.cases$tests.mov.avg<500,"warning","info"))
```


Row
------------
### **Time series of tests**

```{r}
ntl.cases%>%select(c(Date,"New cases"=n.cases,"Tests"=n.tests))%>%pivot_longer(cols=c(2,3))%>%
ggplot( aes(x = Date, y = value,fill=name)) +
  geom_col(size=0.5,position="dodge") +
  #geom_point(size=0.5)+
  scale_x_date(date_breaks = "2 weeks") +
  scale_y_continuous()+
  scale_fill_viridis_d(name="Variable",labels=c("Number of cases","Number of tests"))+
  labs(y = "Daily # of Tests and cases") + 
  theme(panel.background=element_blank(),axis.text.x = element_text(angle = 45, hjust = 1)) -> tests.ts
ggplotly(tests.ts)



```


### **Daily positivity rate and 5 days moving average**
```{r}


ntl.cases%>%select(c(Date,"positivity"=positivity,"Moving.average"=positivity.mov.avg))%>%
  #pivot_longer(cols=c(2,3))%>%
ggplot( aes(x = Date)) +
  geom_col(aes(y=positivity,fill="positivity"),size=0.5,alpha=0.7) +
  geom_line(aes(y= Moving.average,color="Moving.average"),size=1)+
  scale_x_date(date_breaks = "2 weeks",limits = c(as.Date("2020-05-18"),Sys.Date())) +
  labs(y = "Daily positivity (%)")+
  scale_fill_manual(name = "", values = c("positivity" = "chartreuse4"),labels="daily positivity") +
  scale_color_manual(name = "", values = c("Moving.average" = "red4")) +

  theme(panel.background=element_blank(),axis.text.x = element_text(angle = 45, hjust = 1)) -> positivity.ts
ggplotly(positivity.ts)


```



Row 
-----------------------------------------------------------------------

### **Total deaths**
```{r}

valueBox(value=sum(cases_by_region$deaths,na.rm=TRUE), caption="Number of deaths caused by COVID-19 reported in Madagascar since 20 March 2020",color="danger",icon="fa-heartbeat")
```

### **Overall case fatality rate**
```{r}

valueBox(value=paste(round(100*sum(cases_by_region$deaths)/sum(cases_by_region$cases),2),"%"),
         caption="Case Fatality Rate",
         icon="fa-virus",
         color="danger")
```

### **daily deaths**
```{r}

valueBox(value=round(last(ntl.cases$d.mov.avg),2),
         caption="Daily COVID-19 deaths (7 days moving average)",
         icon="fa-bible",
         color=ifelse(last(ntl.cases$d.mov.avg)<1,"warning","danger"))
```

Row {data-width=650}
----------------------------
### **Age and Sex of deceased**
```{r}

deaths%>%filter(!is.na(Sex))%>%
  ggplot()+
  geom_histogram(aes(x=Age),position="dodge",binwidth=5,fill="#7570B3")+labs(y="number of deceased individuals",caption="The age and sex of some individuals were not specified during the press briefings and are not presented on this graph") +theme_bw()+facet_grid(~Sex)+scale_y_continuous(limits=c(0,5))->death_hist
death_hist

```

### ***How many people have died on this day and average daily death rate?***
```{r}
ntl.cases%>%rename("deaths"=n.deaths,"moving.average"=d.mov.avg)%>%
  ggplot(aes(x=Date))+
  geom_col(aes(y=deaths,fill="deaths"))+
  geom_line(aes(y=moving.average,color="moving.average"),size=1)+
  labs(y="number of deceased individuals per day")+
  scale_x_date(date_breaks = "2 weeks") +
  scale_y_continuous(limits=c(0,5))+
  scale_fill_manual(name = "", values = c("deaths" = "#1f78b4"),labels="number of deaths") +
  scale_color_manual(name = "", values = c("moving.average" = "#fdbf6f")) +
  theme(panel.background=element_blank(),axis.text.x = element_text(angle = 45, hjust = 1))->death_time
ggplotly(death_time)

```




