---
title: "COVID-19 | Madagascar"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    social: [ "twitter", "facebook"]
    vertical_layout: scroll
    theme: bootstrap
    source_code: https://github.com/fidyras/madadashboard
    navbar:
      - { title: "<img style=\"width: 16px;\" src=\"www/us16.png\" />   English", href: "dashboard_EN.html"} 
      - { title: "<img style=\"width: 16px;\" src=\"www/mada16.png\" />   Malagasy", href: "index.html"}
      - { title: "<img style=\"width: 16px;\" src=\"www/france16.png\" />   French", href: "dashboard_fr.html"}
params:
  lang: "MDG"
---

<!-- Google Tag Manager -->
<script>(function(w,d,s,l,i){w[l]=w[l]||[];w[l].push({'gtm.start':
new Date().getTime(),event:'gtm.js'});var f=d.getElementsByTagName(s)[0],
j=d.createElement(s),dl=l!='dataLayer'?'&l='+l:'';j.async=true;j.src=
'https://www.googletagmanager.com/gtm.js?id='+i+dl;f.parentNode.insertBefore(j,f);
})(window,document,'script','dataLayer','GTM-N9D53DP');</script>
<!-- End Google Tag Manager -->

```{r setup, include=FALSE}
library(flexdashboard)
library(googlesheets4)
library(shiny)
library(sf)
library(leaflet)
library(tidyverse)
library(plotly)
library(lubridate)
library(dygraphs)
library(zoo)
library(forecast)
library(stringdist)
library(glue)
library(ggthemes)
library(scales)
library(RColorBrewer)
require(EpiEstim)
require(dplyr)
require(RCurl)
require(reshape2)
require(purrr)
require(lubridate)
library(crosstalk)
library(jsonlite)
```


```{r}
# pull data from google sheets & clean
gs4_deauth()
data   <-
  read_sheet(
    "https://docs.google.com/spreadsheets/d/1oQJl4HiTviKAAhCjMmg0ipGcO79cZg6gSHrdTuQID_w/edit#gid=0",
    sheet = 2
  )
test   <-
  read_sheet(
    "https://docs.google.com/spreadsheets/d/1oQJl4HiTviKAAhCjMmg0ipGcO79cZg6gSHrdTuQID_w/edit#gid=0",
    sheet = 3
  )

text   <-
  read_sheet(
    "https://docs.google.com/spreadsheets/d/1oQJl4HiTviKAAhCjMmg0ipGcO79cZg6gSHrdTuQID_w/edit#gid=0",
    sheet = 6
  )
reg_cols <- 
  read_sheet(
    "https://docs.google.com/spreadsheets/d/1oQJl4HiTviKAAhCjMmg0ipGcO79cZg6gSHrdTuQID_w/edit#gid=0",
    sheet = 4
  )

events <-
  read_sheet(
    "https://docs.google.com/spreadsheets/d/1oQJl4HiTviKAAhCjMmg0ipGcO79cZg6gSHrdTuQID_w/edit#gid=0",
    sheet = 5
  )

cases <-
  read_sheet(
    "https://docs.google.com/spreadsheets/d/1oQJl4HiTviKAAhCjMmg0ipGcO79cZg6gSHrdTuQID_w/edit#gid=0",
    sheet = 8
  )

cases_region <-
  read_sheet(
    "https://docs.google.com/spreadsheets/d/1oQJl4HiTviKAAhCjMmg0ipGcO79cZg6gSHrdTuQID_w/edit#gid=0",
    sheet = 9
  )


cases$Date<-as.Date(cases$week)
all.dates<-data.frame(list(Date=seq(min(cases$Date),max(cases$Date),by="day")))
cases<-merge(all.dates, cases, all=T)
cases$n.cases[which(is.na(cases$n.cases))]<-0
cases$n.deaths[which(is.na(cases$n.deaths))]<-0
cases%>%arrange(Date)%>%select(Date,"cases"=n.cases,"deaths"=n.deaths,"tests"=n.tests,"total.tests"=Global.total.test)%>%
  mutate(cases.cumsum = cumsum(cases),
         deaths.cumsum = cumsum(deaths),
         positivity=round((cases/tests),4)) %>% #compute cumulative number of cases per region
  mutate(cases.mov.avg = rollmeanr(cases, 7, fill = "extend"),
         deaths.mov.avg=rollmeanr(deaths, 7, fill = "extend"),
         positivity.mov.avg=rollmeanr(positivity, 7, fill = NA))->cases

# shapefiles
mdg2 <- read_sf("data/regions/MDG_ADM1.shp")

# Cases by region
cases_region$week<-as.Date(cases_region$week)
cases_region$region[is.na(cases_region$region)] = "unspecified"
cases_region%>%group_by(region)%>%summarise(cases=sum(cases),deaths=sum(death,na.rm=TRUE))->region_cumulative #sum of cases by region since March
cases_region$death[which(is.na(cases_region$death))]<-0
cases_region$region <- mdg2$NAME[amatch(cases_region$region, mdg2$NAME, maxDist = 6)]
cases_region$region[is.na(cases_region$region)] = "unspecified"
cases_region%>%
  rename("Date"=week)%>%
  arrange(Date)%>%
  group_by(region) %>%
  mutate(cases.cumsum = cumsum(cases),
         deaths.cumsum=cumsum(death)) %>% #compute cumulative number of cases per region
  mutate(cases.mov.avg = rollmeanr(cases, 7, fill = "extend"),
         death.mov.avg = rollmeanr(death, 7, fill = "extend"))%>%ungroup()-> cases_region #compute 7 day moving average per region



```


``` {r translation}
# Seting up language for rest of dashboard
# Selecting only the column with the text corresponding to the selected language
c.lang <- params$lang
text <- text %>% select("ref", all_of(c.lang))
# Writing a helper function to pull the language specified in params above
f.tr <- function(ref) {
  txt <- as.character(text[text$ref == ref, 2])
  return(txt)
}
```


`r f.tr("weekly")`
=======================================================================

Sidebar {.sidebar}
-----------------------------------------------------------------------
\

```{r, results='asis', verbatim=TRUE}
date_updated <- format(last(cases$Date), format = "%B %d %Y") # any inline stats you need ahead of time!
```

`r f.tr("sidebar_part_1")`

`r f.tr("sidebar_part_2")`

`r f.tr("sidebar_part_3")`

`r f.tr("sidebar_part_4")`

`r paste(f.tr("sidebar_part_5"), date_updated, sep = " ")` 

`r f.tr("sidebar_part_7")`

`r f.tr("sidebar_part_8")`

`r f.tr("sidebar_part_9")`

Row 
-----------------------------------------------------------------------

### cases

```{r}
valueBox(
  value = format(sum(cases$cases,na.rm=TRUE), big.mark = ","),
  caption = paste("total number of cases until",
                  format(last(cases$Date), format = "%B %d %Y")),
  icon = "fa-ambulance",
  color = "info"
)
```

### deaths

```{r}

valueBox(
  value = sum(cases$deaths, na.rm = TRUE),
  caption = paste("total number of deaths recorded until",
                  format(last(cases$Date), format = "%B %d %Y")),
  color = "danger",
  icon = "fa-heartbeat"
)

```

### cases

```{r}

valueBox(
  value = format(last(cases$cases), big.mark = ","),
  caption = paste("Number of cases reported on",
    format(last(cases$Date), format = "%B %d %Y")),
  icon = "fa-viruses",
  color = "warning"
  )


```

### tests

```{r}

valueBox(
  value = format(last(cases$tests), big.mark = ","),
  caption = paste("Number of tests performed on",
    format(last(cases$Date), format = "%B %d %Y")),
  icon = "fa-flask",
  color ="info")


```



### Positivity

```{r}


gauge(
  round(last(cases$positivity)*100, 2),
  min = 0,
  max = 100,
  symbol = '%',
  label = paste("Positivity on \n",
    last(cases$Date)),
  abbreviateDecimals = 2,
  sectors = gaugeSectors(
    success = c(0, 5),
    warning = c(5, 20),
    danger = c(20, 100)
  )
)

```





Row 
-----------------------------------------------------------------------
### Number of new cases

```{r, warning=FALSE}

daily.cases <- ggplot(data = cases, aes(x = Date, y = cases.mov.avg)) +
  geom_col(aes(x = Date, y = cases, fill = "cases",
               text = map(paste(f.tr("time_series_c_xlab"), ":", Date, "<br>",
                                f.tr("time_series_c_ylab"), ":", cases, "<br>",
                                f.tr("times_series_c_textlab"), ":", round(cases.mov.avg, 2), "<br>"), HTML)),
           position = "dodge",
           alpha = 0.4) +
  scale_x_date(date_breaks = "2 weeks", date_labels = "%Y-%m-%d") +
  geom_line(aes(x = Date, y = cases.mov.avg, color = "cases.mov.avg"),
            size = 1) +
  scale_fill_manual(name = "", values = c("cases" = "#1B9E77")) +
  scale_color_manual(name = "", values = c("cases.mov.avg" = "#0d4f3b")) + # matched moving avg cols to bar cols
  theme(legend.position = "none", 
        panel.background = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = f.tr("time_series_c_xlab"),
       y = f.tr("time_series_c_ylab"))
ggplotly(daily.cases, tooltip = "text") %>% 
  layout(hovermode = "x") %>%
  config(displaylogo = FALSE)
```


Row 
-----------------------------------------------------------------------
### `r f.tr("map_of_cases_header")`

`r f.tr("map_of_cases_note")`

```{r}
mada_regions <-
  left_join(mdg2, region_cumulative, by = c("NAME" = "region"))
bins <- round(exp(pretty(log(region_cumulative$cases), n = 10)), 0) # logged breaks
pal <- colorBin("YlOrRd", domain = mada_regions$cases, bins = bins)
labels <- sprintf(
  "<strong>%s</strong><br/> Cases: %i <br/>",
  mada_regions$NAME,
  mada_regions$cases
) %>% lapply(htmltools::HTML)
map <- leaflet()  %>%
  addTiles() %>%
  addPolygons(
    data = mada_regions,
    color = "black",
    weight = 0.001,
    smoothFactor = 0.1,
    fillColor = ~ pal(cases),
    fillOpacity = 0.6,
    dashArray = NULL,
    label = labels,
    highlightOptions = highlightOptions(
      weight = 3,
      color = "black",
      dashArray = NULL,
      fillOpacity = 0.75,
      bringToFront = TRUE
    ),
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto"
    )
  ) %>%
  addLegend(
    "topleft",
    pal = pal,
    values = bins,
    title = f.tr("map_of_cases_legend_caption")
  )
map %>% addProviderTiles(providers$Esri.WorldStreetMap)
```


```{r, warning=FALSE}
# This code uses the library crosstalk to link the graphs
# Colors pulled from google sheets (please change as needed!)
mycolors <- reg_cols$color
names(mycolors) <- reg_cols$shapefile_name
mycolors <- c(mycolors, c("unspecified" = "grey"))
# Order by cumulative sums
cases_region %>%
  group_by(region) %>%
  mutate(max_sum = max(cases.cumsum)) %>%
  ungroup() %>% 
  mutate(region = reorder(region, -max_sum)) -> cases_by_week
cases_by_week$cheat <- "Reset to all regions when no selection"
sd <- SharedData$new(cases_by_week, ~region, group = "Choose region")
sd %>%
  ggplot(aes(x = Date, y = cum.sum, fill = region, 
           text = map(paste("<b>", region, "</b><br>",
                       f.tr("time_series_b_xlab"), ":", Date, "<br>",
                       f.tr("time_series_b_ylab"), ":", cases.cumsum, "<br>"), HTML))) +
  geom_area(size = 0.75) +
  scale_fill_manual(values = mycolors) + 
  scale_x_date(date_breaks = "2 weeks", date_labels = "%Y-%m-%d") +
  theme(panel.background = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = f.tr("time_series_b_xlab"),
       y = f.tr("time_series_b_ylab")) -> cumulative_cases

sd%>%
  ggplot() +
  geom_col(
    aes(x =Date, y = cases, fill = region),
    position = "dodge",
    alpha = 0.2,
    size = 0.1
  ) +
  scale_x_date(date_breaks = "2 weeks", date_labels = "%Y-%m-%d") +
  scale_color_manual(values = mycolors, name = "Region") +
  scale_fill_manual(values = mycolors, name = "Region") +
  labs(x = f.tr("time_series_d_xlab"),
       y = f.tr("time_series_d_ylab")) +
  theme(panel.background = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1)) -> modavg.curv
```


### `r format(f.tr("time_series_b_header"))` & `r format(f.tr("time_series_c_header"))`

```{r}
bscols(
  filter_select(
      id = "region",
      label = "Region",
      sharedData = sd,
      group = ~region,
      multiple = TRUE,
  ),
  filter_checkbox(id = "regall", label = "", sharedData = sd, group = ~cheat)
)
bscols(
  (ggplotly(modavg.curv, tooltip = "text",dynamicTicks=TRUE) %>% 
    layout(yaxis = list(autorange = TRUE)) %>%
    highlight(on = NULL, off = NULL) %>% # Turn off highlighting otherwise a pain
    config(displaylogo = FALSE))
)


```


Row 
-----------------------------------------------------------------------

### Number of deaths

```{r, warning=FALSE}
cases %>%
  ggplot(aes(x = Date, 
             text =  map(paste(f.tr("mortality_t_xlab"), Date, "<br>",
                               f.tr("mortality_t_ylab"), ":", deaths, "<br>",
                               f.tr("mov_average"), ":", round(deaths.mov.avg, 2), "<br>"), HTML))) +
  geom_col(aes(y = deaths, fill = "deaths")) +
  geom_line(aes(y = deaths.mov.avg, color = "deaths.mov.avg"), size = 1) +
  labs(x = f.tr("mortality_t_xlab"),
       y = f.tr("mortality_t_ylab")) +
  scale_x_date(date_breaks = "2 weeks") +
  scale_y_continuous(limits = c(0, 15)) +
  scale_fill_manual(
    name = "",
    values = c("deaths" = "#ff6262"),
    labels = "number of deaths"
  ) +
  scale_color_manual(name = "",
                     values = c("deaths.mov.avg" = "#9d0000")) +
  theme(
    legend.position = "none",
    panel.background = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) -> death_time
ggplotly(death_time, tooltip = "text")

```
Row 
-----------------------------------------------------------------------

### Case Fatality Ratio

Through `r last(cases$Date)`, Madagascar has officially registered `r format(last(cases$cases.cumsum),scientific=FALSE)` cases and `r format(last(cases$deaths.cumsum),scientific=FALSE)` deaths.



Using data on deaths that occurred and the number of cases, we can estimate the **case fatality ratio (CFR)**. The CFR is defined as the proportion of **confirmed COVID-19 deaths** over the number of **confirmed COVID19 cases**.



The CFR can help us understand the severity of the disease in confirmed cases however it is *not the same as* the risk of death for an infected person and thus **does not** represent the probability of dying if an individual is infected by COVID-19.




In addition, *the final clinical outcome of most of the reported cases is unknown*. Therefore the tally of deaths over the confirmed cases might underestimate the true case fatality ratio.

### case fatality ratio

```{r}

valueBox(
  value = paste(round(100 * sum(cases$deaths) / sum(cases$cases), 2), "%"),
  caption = f.tr("cfr_caption"),
  icon = "fa-virus",
  color = "danger"
)

```



Row 
-----------------------------------------------------------------------
### Case Fatality Ratio

```{r cfr}
cases<-cases%>%
  mutate(cfr=round(deaths.cumsum/cases.cumsum*100,4))



cases_region%>%
  filter(!is.na(region),region!="unspecified")%>%
  group_by(region)%>%
  mutate(cum.deaths=cumsum(death),cum.cases=cumsum(cases),cfr=round(cum.deaths/cum.cases,4)*100)%>%
  ggplot()+
  geom_line(aes(x=Date,y=cfr,color=region,group=region),alpha=0.2)+
  geom_line(data=cases,aes(x = Date, y = cfr),method="gam",se=FALSE) +
  scale_y_continuous(name="case fatality rate (%)",limits = c(0,7), breaks=seq(0,10,0.5))+
  scale_x_date(date_breaks = "2 weeks")+
  scale_color_manual(values = mycolors, name = "Region")+
  theme(legend.position = "none", 
        panel.background = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Week ending on")->cfr_reg_plot
ggplotly(cfr_reg_plot)%>%
  #layout(hovermode = "x") %>%
  config(displaylogo = FALSE)


  

```





Row 
-----------------------------------------------------------------------

### Number of tests


```{r}
cases %>% 
  select(c(Date, "New cases" = cases, "Tests" = tests)) %>%
  mutate(minus_pos = Tests - `New cases`, 
         Tests = ifelse(is.na(Tests), `New cases`,  minus_pos),
         total_tests = Tests + `New cases`) %>%
  select(-minus_pos) %>%
  pivot_longer(2:3) %>%
  mutate(name = fct_relevel(name, "New cases", after = Inf), 
         label = case_when(name == "New cases" ~ value, 
                           name == "Tests" ~ total_tests),
         name = case_when(name == "New cases" ~ f.tr("tests_t_lab_cases"),
                          name == "Tests" ~ f.tr("tests_t_lab_tests"))) -> tests_pos
test_cols <- c("#8d66a1", "#cdbbd5")
ggplot(data = tests_pos, 
       aes(x = Date, y = value, fill = name, 
           text = map(paste(f.tr("tests_t_xlab"), ":", Date, "<br>",
                             name, ":", label, "<br>"), HTML))) +
  geom_col(size = 0.5, position = "stack") +
  # geom_point(size = 0.5) +
  scale_x_date(date_breaks = "2 weeks", date_labels = "%Y-%m-%d") +
  scale_fill_manual(values = test_cols, name = "") +
  labs(x = f.tr("tests_t_xlab"), y = f.tr("tests_t_ylab")) +
  theme(panel.background = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1)) -> tests.ts
ggplotly(tests.ts, tooltip = "text") %>% 
  layout(hovermode = "x") %>% 
  config(displaylogo = FALSE)

```



### Positivity


```{r}
cases%>%
  ggplot(aes(x = Date, 
             text =  map(paste(f.tr("daily_pos_xlab"), Date, "<br>",
                               f.tr("daily_pos_ylab"), ":", round(positivity, 2), "<br>",
                               f.tr("mov_average"), ":", round(positivity.mov.avg, 2), "<br>"), HTML))) +
  geom_col(aes(y = positivity, fill = "positivity"),
           size = 0.5,
           alpha = 0.7) +
  geom_line(aes(y = positivity.mov.avg, color = "positivity.mov.avg"), size = 1) +
  scale_y_continuous(labels=percent)+
  scale_x_date(date_breaks = "2 weeks",
               limits = c(as.Date("2020-05-18"), Sys.Date())) +
  labs(x = f.tr("daily_pos_xlab"),
       y = f.tr("daily_pos_ylab")) +
  scale_fill_manual(
    name = "",
    values = c("positivity" = "#00a579"),
    labels = "daily positivity"
  ) +
  scale_color_manual(name = "", values = c("positivity.mov.avg" = "#004331")) +
  theme(
    legend.position = "none",
    panel.background = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) -> positivity.ts
ggplotly(positivity.ts, tooltip = "text")

```


R~t~ estimates
===================================

**Estimating R~t~ for Madagascar** {.sidebar}
-----------------------------------------------------------------------

```{r}
rt_summary <-
  data.table::fread(
    "https://raw.githubusercontent.com/labmetcalf/mada-rt/master/latest/rt_summary.csv"
  )
rt_ests <-
  data.table::fread(
    "https://raw.githubusercontent.com/labmetcalf/mada-rt/master/latest/rt_ests.csv"
  )
nowcast <- 
  data.table::fread(
    "https://raw.githubusercontent.com/labmetcalf/mada-rt/master/latest/case_cast.csv"
  )
```
\
\
The time varying reproductive number (R~t~), which represents the average number of secondary infections generated by each new infectious case, is a way to track the progress of an outbreak. 
\
\
If R~t~ in the population is greater than 1, cases will continue to grow in number. If R~t~ is less than 1, case numbers will decline. The higher the value of R~t~ above 1, the faster an epidemic will grow. 
\
\
R~t~ was estimated using the using the R packages [`epiNow`](https://epiforecasts.io/EpiNow/), [`epiSoon`](https://github.com/epiforecasts/EpiSoon), and [`epiEstim`](https://cran.r-project.org/web/packages/EpiEstim/index.html), code available [here](https://github.com/labmetcalf/mada-rt).
\
\
These estimates were last updated on `r format(min(ymd(rt_summary$date_ests)), "%Y-%m-%d")`

Row 
-----------------------------------------------------------------------

### **R~t~ National Estimate** 

```{r}
# Updating so that it is the latest nowcast
rt_summary %>%
  filter(Region == "National") %>% .$`Effective reproduction no.` -> rt_national
gauge(rt_national, min = 0, max = 5, symbol = '',
      label = paste("Nowcast on", format(rt_summary$date[1], format="%Y-%m-%d")),
      abbreviateDecimals = 2,
      sectors = gaugeSectors(success=c(0,1), warning = c(1, 2), danger = c(2, 5)))
```

### Expected cases

```{r}
rt_summary %>%
  filter(Region == "National") %>%
  .$`Expected change in daily cases` -> expected
box_col <- case_when(expected == "Increasing" ~ "danger", 
                     expected == "Likely increasing" ~ "LightCoral", 
                     expected == "Unsure" ~ "Grey", 
                     expected == "Likely decreasing" ~ "DarkSeaGreen", 
                     expected == "Decreasing" ~ "DarkOliveGreen")
valueBox(expected, 
         caption = "Expected change in daily cases",
         color = box_col)
```

Row {.tabset data-height=800}
-----------------------------------------------------------------------

```{r helper}
# Helper to autogenerate tabs
gen_rt_tabs <- function(region, rt_ests, nowcast, minrt = 0) {
  
  mycolors <- c(mycolors, National = "navy")
  cols <- c(alpha(mycolors[names(mycolors) == region], 0.9), 
            alpha(mycolors[names(mycolors) == region], 0.7), 
            alpha(mycolors[names(mycolors) == region], 0.4))
  
  names(cols) <- c("Observed by report date", "nowcast", "forecast")
  ggplot(filter(rt_ests, region == !!region, date >= "2021-03-15"), 
         aes(x = as.Date(date), y = median, group = rt_type, fill = rt_type)) +
    geom_ribbon(aes(ymin = bottom, ymax = top), alpha = 0.4) +
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.7) +
    geom_hline(yintercept = 1, linetype = 3) +
    labs(x = "Date", y = "Effective reproductive number") +
    scale_x_date(date_breaks = "2 weeks", date_labels = "%Y-%m-%d") +
    scale_linetype_manual(values = c(2, 1), name = "") +
    scale_fill_manual(values = cols, name = "") +
    expand_limits(y=0) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) -> rtcast
  
  # Manually fix legends
  rp <- ggplotly(rtcast)
  rp[['x']][['data']][[1]][['name']] <- "R<sub>t</sub> Forecast"
  rp[['x']][['data']][[2]][['name']] <- 'R<sub>t</sub> Nowcast'
  
  
  ggplot(filter(nowcast, region == !!region, date >= "2021-03-15"), 
         aes(x = ymd(date), group = type, fill = type)) +
    geom_col(data = filter(nowcast, region == !!region, date >= "2021-03-15", 
                      type == "Observed by report date"),
             aes(y = median, fill = type), alpha = 0.8, position = "dodge") +
    geom_col(data = filter(nowcast, region == !!region, date >= "2021-03-15", 
                      type != "Observed by report date"),
             aes(y = median, fill = type), alpha = 0.8, position = "dodge") + 
    geom_ribbon(aes(ymin = bottom, ymax = top), alpha = 0.4) +
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.7) +
    scale_fill_manual(values = cols, name = "") +
    scale_x_date(date_breaks = "2 weeks", date_labels = "%Y-%m-%d") +
    guides(fill = guide_legend(nrow = 2, byrow = TRUE)) +
    labs(x = "Date", y = "Cases") +
    theme(legend.position = "bottom", 
          axis.text.x = element_text(angle = 45, hjust = 1)) -> now
  
  # Manually fix legends
  cp <- ggplotly(now)
  cp[['x']][['data']][[1]][['name']] <- 'Cases by confirmation \n date (observed)'
  cp[['x']][['data']][[2]][['name']] <- 'Case Forecast'
  cp[['x']][['data']][[3]][['name']] <- 'Case Nowcast'
  return(
    subplot(
    style(ggplotly(rp), hoverinfo = "none"),
    style(ggplotly(cp), hoverinfo = "none"),
    nrows = 2, 
    shareX = TRUE, shareY = FALSE
  )
  )
}
  
```

```{r autotabs, results = 'asis'}
# order by case #s
nowcast %>% 
  group_by(region) %>% 
  filter(type == "Observed by report date") %>% 
  summarize(cases = sum(median)) %>% 
  arrange(desc(cases)) -> regs

for(i in 1:length(regs$region)) {
  p <- gen_rt_tabs(region = regs$region[i], rt_ests, nowcast)
  cat(paste("###", regs$region[i]), "\n")
  cat(as.character(htmltools::tagList(p)))
}
```


```


Row {data-height=400}
-------------------------------

### Notes

* We simulated infection dates (n = 500) using the methods described  [here](https://wellcomeopenresearch.org/articles/5-112/v1).

* Expected change in daily cases were classified as:  
    + Increasing: < 5% of resulting R~t~ estimates were below 1.
    + Likely increasing: < 20% of resulting R~t~ estimates were below 1.
    + Unsure: 20 - 80% of resulting R~t~ estimates were below 1
    + Likely decreasing: > 80% of resulting R~t~ estimates were below 1
    + Decreasing: > 95% of of resulting R~t~ estimates were below 1
    
* For model estimates of R~t~ and cases, the darker shading shows the 50% credible intervals and the lighter shaded ribbon shows the 90% credible intervals.

* Forecasts of R~t~ at low/zero case counts have wide credible intervals and low confidence.

* There are a number of limitations and uncertainties in these methods, many of which are reviewed [here](https://www.medrxiv.org/content/10.1101/2020.06.18.20134858v2?rss=1). 


```{r comparison}
# starting code for comparing Mada to other countries
# following instructions from here:
# https://stackoverflow.com/questions/50161492/how-do-i-scrape-data-from-an-arcgis-online-map
# pull in Africa cdc data
afr_cdc <- jsonlite::fromJSON("https://services8.arcgis.com/vWozsma9VzGndzx7/arcgis/rest/services/NEW_Dashboard/FeatureServer/0/query?where=0%3D0&outFields=%2A&f=json")
afr_cdc_df <- tibble::as_tibble(afr_cdc$features$attributes)
# add in Mada data and compare
# ggplot(data = filter(afr_cdc_df, !is.na(Tests)), 
#        aes(x = reorder(Country, Tests/`2020_Population`*1e5), 
#            y = Tests/`2020_Population`*1e5)) +
#   geom_col() +
#   labs(x = "", y = "Tests per 100k persons") +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
```