---
title: "Covid-19 in Madagascar blabla"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    theme: bootstrap
---

```{r setup, include=FALSE}
library(flexdashboard)
library(googlesheets4)
library(shiny)
library(sf)
library(leaflet)
library(tidyverse)
library(plotly)
library(lubridate)
library(dygraphs)
library(zoo)
library(forecast)
library(stringdist)
```


```{r}

# pull data from google sheets & clean
gs4_deauth()
data <- read_sheet("https://docs.google.com/spreadsheets/d/1oQJl4HiTviKAAhCjMmg0ipGcO79cZg6gSHrdTuQID_w/edit#gid=0", sheet = 5)
test <- read_sheet("https://docs.google.com/spreadsheets/d/1oQJl4HiTviKAAhCjMmg0ipGcO79cZg6gSHrdTuQID_w/edit#gid=0", sheet = 6)
events <- read_sheet("https://docs.google.com/spreadsheets/d/1oQJl4HiTviKAAhCjMmg0ipGcO79cZg6gSHrdTuQID_w/edit#gid=0", sheet = 8)
# shapefiles
mdg2 <- read_sf("data/regions/MDG_ADM1.shp")

# Cases by region
data$region<-mdg2$NAME[amatch(data$Location4,mdg2$NAME,maxDist=6)]
data$region[is.na(data$region)]="unspecified"
data %>%
  group_by(region) %>%
  summarize(cases = sum(Type=="N"), deaths = sum(Type == "D")) -> cases_by_region


deaths<-filter(data,Type=="D")
cases<-filter(data,Type=="N")

cases_by_date <- as.data.frame(xtabs( ~ Date+region , cases))
cases_by_date$Date<-as.Date(cases_by_date$Date)
cases_by_date%>%rename(cases=Freq)%>%
  group_by(region)%>%
  mutate(cum.sum=cumsum(cases))%>%#compute cumulative number of cases per region
  mutate(mov.avg=ma(cases,order=7))->cases_by_date#compute 7 days moving average per region
deaths$Age<-as.numeric(deaths$Age)
deaths_by_date<-as.data.frame(xtabs( ~ Date+region , deaths))
deaths_by_date$Date<-as.Date(deaths_by_date$Date)
deaths_by_date%>%
  rename(deaths=Freq)%>%
  group_by(region)%>%
  mutate(deaths.cum.sum=cumsum(deaths))%>%#compute cumulative number of cases per region
  mutate(deaths.mov.avg=ma(deaths,order=7))%>% #compute 7 days moving average
  right_join(cases_by_date)->cases_by_date
#IRD for infected (new cases), Recovered (R) and Removed (D)
IRD<-as.data.frame(xtabs( ~ data$Date+data$Type))
IRD%>%rename("Date"=1,"Type"=2,"n"=3)%>%filter(Type!="S")->IRD
IRD$Date<-as.Date(IRD$Date)

IRD%>%
  arrange(Date)%>%
  dplyr::group_by(Type)%>%
  mutate(Type.cum.sum=cumsum(n))->IRD


##testing

test %>%select("Date"=Date,"Cum.cases"=Cum.cases,"New.cases"=new.cases, "Total.tests"=Global.total.test)->test

Diff.previous<-data.frame("Date"=test$Date[c(2:length(test$Total.tests))],
                          "tests.diff"=diff(test$Total.tests, lag = 1,na.rm=T),
                                               "time.diff"=as.integer(diff.Date(test$Date)))%>%
  filter(!is.na(tests.diff)&time.diff==1)#table of daily test (when  there is exactly 1 day between two available "numbers of tests")

cases%>%group_by(Date)%>%
  summarise(n.cases=sum(Type=="N"))%>%
  mutate(cum.cases=cumsum(n.cases),mov.avg=ma(n.cases,order=7))%>%
  right_join(test)%>%
  left_join(Diff.previous)%>%
  mutate(tests.mov.avg=ma(tests.diff,order=7))%>%
  mutate(positivity=(n.cases/tests.diff)*100)%>%
  mutate(positivity.mov.avg=ma(positivity,order=7))%>%
  select(Date,n.cases,cum.cases,mov.avg,n.tests=tests.diff,cum.tests=Total.tests,tests.mov.avg,positivity,positivity.mov.avg)->ntl.cases

ntl.cases<-ntl.cases[order(ntl.cases$Date,na.last=FALSE),]
ntl.cases$cum.cases<-na.locf(ntl.cases$cum.cases)
ntl.cases$Date<-as.Date(ntl.cases$Date)

events$Date<-as.Date(events$`Start date`)
events$event<-events$`Events/Measures`

```
Country profile: Madagascar
========================================================================

Sidebar {.sidebar}
-----------------------------------------------------------------------
This country profile of Madagascar facing the COVID-19 pandemic includes interactive visualizations and limited explanations of the presented metrics.


Data are obtained from daily press briefings presented by the CoronaVirus Operational Task force of the Government of Madagascar.




The details and accuracy of the figures presented here are contingent upon the publicly available information communicated to the public by the Government of Madagascar.








The data on the coronavirus cases presented on this dashboard are updated daily. 
Last update: 
```{r}
format(last(cases$Date),format="%B %d %Y")
```




You can download our complete dataset here and the R code to all figures here





Links to Versions in French and Malagasy.



Row 
-----------------------------------------------------------------------

### Total cases

```{r}
valueBox(value=format(sum(cases_by_date$cases),big.mark = ","),
         caption=paste("Number of cases reported in Madagascar since 20-March-2020 until", format(last(cases_by_date$Date), format="%B %d %Y")),
         icon="fa-ambulance",color="info")
```

### Total Recovered
```{r}
IRD%>%filter(Type=="R")->recov
valueBox(value=format(last(recov$Type.cum.sum),big.mark = ","),
         caption=paste("Number of recoveries reported in Madagascar since 20-March-2020 until", format(last(recov$Date), format="%B %d %Y")),
         icon="fa-virus-slash",color="green")
```

### Total deaths
```{r}
valueBox(value=sum(cases_by_date$deaths,na.rm=TRUE), 
         caption="Total Number of deaths caused by COVID-19 reported in Madagascar since 20 March 2020",color="danger",icon="fa-heartbeat")
```


### Total tests performed: How many tests did Madagascar perform so far?
```{r}
valueBox(value=format(last(test$Total.tests),big.mark = ","),
         caption=paste("Total number of tests in Madagascar since 20-March-2020 until", format(last(test$Date), format="%B %d %Y")),
         icon="fa-flask",color=ifelse(test$Total.tests/27692<5,"warning","info"))
```


### Positivity (7-day moving avergage)
```{r}
positivity_ma<-ntl.cases%>%filter(!is.na(positivity.mov.avg))
gauge(round(last(positivity_ma$positivity.mov.avg),2),
      min=0,max=100,symbol='%',
      label="Average positivity rate",abbreviateDecimals = 2,
      sectors=gaugeSectors(success=c(0,5),warning = c(5,20),danger=c(20,100)))
```







Row {data-width=650}
-----------------------------------------------------------------------

### Map of cases 

```{r}
mada_regions <- left_join(mdg2, cases_by_region, by = c("NAME" = "region"))

bins <- pretty(cases_by_region$cases)
pal <- colorNumeric("YlOrRd", domain = mada_regions$cases)

labels <- sprintf(
      "<strong>%s</strong><br/> Cases: %i <br/> Deaths: %i",
      mada_regions$NAME, mada_regions$cases, mada_regions$deaths
    ) %>% lapply(htmltools::HTML)
    
m<-leaflet()  %>%
  addTiles()%>%
  addPolygons(data = mada_regions,
              color = "black", weight = 0.001, smoothFactor = 0.1,
              fillColor = ~pal(cases),
              fillOpacity = 0.6,
              dashArray = NULL,
              label = labels,
              highlightOptions = highlightOptions(
                    weight = 3,
                    color = "black",
                    dashArray = NULL,
                    fillOpacity = 0.75,
                    bringToFront = TRUE),
              labelOptions = labelOptions(
                style = list("font-weight" = "normal", padding = "3px 8px"),
                textsize = "15px",
                direction = "auto")) %>%
  addLegend("bottomright", pal = pal, values = bins, title = "Number of reported cases")
m%>%addProviderTiles(providers$Esri.WorldStreetMap)
```

### Cumulative number of cases

```{r}
ggplot(data=cases_by_date,aes(x=Date,y=cum.sum,fill=region))+
  geom_area(data=cases_by_date,aes(x=Date,y=cum.sum,fill=region),size=0.75)+
  scale_fill_brewer(palette="Paired")+scale_x_date(date_breaks = "2 weeks")+theme(panel.background=element_blank(),axis.text.x = element_text(angle = 45, hjust = 1))+labs(title="Cumulative Number of cases reported",y="Number of cases")->cumulative_cases
ggplotly(cumulative_cases)

```



Row {data-width=650}
-----------------------------------------------------------------------


### Time series New cases, Recoveries and deaths

```{r}
ggplot(data=IRD,aes(x=Date,y=Type.cum.sum,fill=Type))+
  geom_area(data=IRD,aes(x=Date,y=Type.cum.sum,fill=Type),alpha=0.7,size=2,position="dodge")+
  geom_line()+
  scale_fill_discrete(labels=c("New cases","Recovered","Deaths"),breaks=c("N","R","D"))+
  scale_x_date(date_breaks = "2 weeks")+
  theme(panel.background=element_blank(),axis.text.x = element_text(angle = 45, hjust = 1))+
  labs(title="New cases, recoveries and deaths",y="Number")->IRD_curve
IRD_curve

```


### Daily number of cases and moving average

```{r, warning=FALSE}
daily.cases<-ggplot(data=ntl.cases,aes(x=Date,y=mov.avg))+
  geom_col(aes(x=Date,y=n.cases),position="dodge",alpha=0.6)+
  scale_x_date(date_breaks = "2 weeks") +
  geom_smooth(aes(x=Date,y=mov.avg,),size=0.7,method="gam")+
  labs(title="Number of cases reported daily and 7-day moving average",subtitle =  "National level",y="Daily number of cases")+
  theme(panel.background=element_blank(),axis.text.x = element_text(angle = 45, hjust = 1))
daily.cases<-daily.cases+geom_vline(aes(xintercept=Date),data=events)
ggplotly(daily.cases)
```

Row {data-width=650}
----------------------------------
### Daily Number of cases per region 
```{r cases_movavg, warning=FALSE}
modavg.curv<-ggplot(data=cases_by_date,aes(x=Date,y=mov.avg))+
  geom_col(aes(x=Date,y=cases,fill=region),position="dodge",alpha=0.6)+
  scale_x_date(date_breaks = "2 weeks") +
  geom_smooth(aes(x=Date,y=mov.avg,group=region,colour=region),size=0.7,method="gam")+
  labs(title="Number of cases reported daily and 7-day moving average",subtitle="Regional level",y="Daily number of cases")+
  theme(panel.background=element_blank(),axis.text.x = element_text(angle = 45, hjust = 1))
ggplotly(modavg.curv)



```

Row 
-----------------------------------------------------------------------

### Total tests performed: How many tests did Madagascar perform so far?
```{r}
valueBox(value=format(last(test$Total.tests),big.mark = ","),caption=paste("Total Number of tests performed in Madagascar since 20-March-2020", Sys.Date()),
         icon="fa-flask",color=ifelse(last(test$Total.tests)/27692<5,"warning","info"))
```

### Tests per thousand: How many tests did Madagascar perform so far?
```{r}
valueBox(value=format(round(last(test$Total.tests/27692),2),big.mark = ","),caption="Tests per thousand habitants performed in Madagascar",
         icon="fa-flask",color=ifelse(last(test$Total.tests)/27.692<5,"warning","info"))
```

### Positivity (7-day moving avergage)
```{r}
positivity_ma<-ntl.cases%>%filter(!is.na(positivity.mov.avg))
gauge(round(last(positivity_ma$positivity.mov.avg),2),min=0,max=100,symbol='%',label="Positivity rate",abbreviateDecimals = 2,
      sectors=gaugeSectors(success=c(0,5),warning = c(5,20),danger=c(20,100)))
```



Row
------------
### Time series of tests 

```{r}
ntl.cases%>%select(c(1,2,5))%>%pivot_longer(cols=c(2,3))%>%
ggplot( aes(x = Date, y = value,fill=name)) +
  geom_col(size=0.5,position="dodge") +
  #geom_point(size=0.5)+
  scale_x_date(date_breaks = "2 weeks") +
  scale_y_continuous()+
  scale_fill_viridis_d(name="Variable",labels=c("Number of cases","Number of tests"))+
  labs(y = "Daily # of Tests and cases") + 
  theme(panel.background=element_blank(),axis.text.x = element_text(angle = 45, hjust = 1)) -> tests.ts
ggplotly(tests.ts)



```
### Trends in Positivity rate and 7 days moving average
```{r}


ntl.cases%>%select(c(1,8,9))%>%pivot_longer(cols=c(2,3))%>%
ggplot( aes(x = Date, y = value,color=name)) +
  geom_line(size=0.7) +
  geom_point(size=0.5)+
  scale_x_date(date_breaks = "2 weeks",limits = c(as.Date("2020-05-18"),Sys.Date())) +
  scale_color_brewer(palette="Dark2")+
  labs(y = "Daily positivity (%)",color="Indicator")+
  theme(panel.background=element_blank(),axis.text.x = element_text(angle = 45, hjust = 1)) -> positivity.ts
ggplotly(positivity.ts)


```



Row 
-----------------------------------------------------------------------

### Total deaths
```{r}

valueBox(value=sum(cases_by_date$deaths,na.rm=TRUE), caption="Number of deaths caused by COVID-19 reported in Madagascar since 20 March 2020",color="danger",icon="fa-heartbeat")
```

### Overall case fatality rate 
```{r}

valueBox(value=paste(round(100*sum(cases_by_region$deaths)/sum(cases_by_region$cases),2),"%"),
         caption="Case Fatality Rate",
         icon="fa-book-medical",
         color="danger")
```

Row {data-width=650}
----------------------------
### Age and Sex of deceased
```{r}

deaths%>%filter(!is.na(Sex))%>%
  ggplot()+
  geom_histogram(aes(x=Age),position="dodge",binwidth=10)+theme_bw()+facet_grid(~Sex)->death_hist
ggplotly(death_hist)
```

### Case Fatality Rate of the ongoing pandemic
```{r}
cases_by_date%>%mutate(CFR=(deaths.cum.sum/cum.sum)*100)%>%
  filter(!is.na(CFR))%>%filter(region!="unspecified")%>%
  ggplot()+
  geom_point(aes(x=Date,y=CFR,group=region,color=region))+
  geom_line(aes(x=Date,y=CFR,group=region,color=region),se=FALSE)+
  scale_x_date(date_breaks = "2 weeks") +
  labs(title="Trends of Case Fatality Rate in Madagascar",y="CFR (%)")+
  theme(panel.background=element_blank(),axis.text.x = element_text(angle = 45, hjust = 1))->CFR
cases_by_date%>%
  group_by(Date)%>%
  mutate(deaths.ntl=sum(deaths.cum.sum,na.rm=TRUE),cases.ntl=sum(cum.sum),CFR.ntl=(deaths.ntl/cases.ntl)*100)->CFR.ntl
CFR+geom_smooth(data=CFR.ntl,aes(x=Date,y=CFR.ntl))+geom_point(data=CFR.ntl,aes(x=Date,y=CFR.ntl),size=0.5)
```




